\LARGE \bf Mean Estimation from One-Bit Measurements

### 2022-05-04 uniform-are-weirdos.tex
finished response and revision
arranged appendix in two-columns

\newtext{Let us now consider the probability that $\theta_n\init$ is substantially wrong. Set $\theta_M \equiv \max_{i \in I_j^*} X_i - 1$. Note that we always have $\theta\init_n \ge \theta_M$ because no observations will be above
$t_j^*+1$, and that $\theta\init_n \le \theta_M + 2n\epsilon/m$. 
 In addition, 
\begin{align*}
P_\theta\left( \left| \theta_M - \theta \right| \ge \frac{2n\epsilon}{m} \right) = \left(1 - \frac{2 n \epsilon}{m} \right)^{n/m}.
\end{align*}
Putting it all together using the triangle inequality, we have
\begin{align*}
  P_\theta\left( \left| \theta\init_n - \theta \right| \ge \frac{4n\epsilon}{m} \right) \le \left(1 - \frac{2 n \epsilon}{m} \right)^{n/m} \le e^{-2 \frac{n^2}{m^2} \epsilon}
  .
\end{align*}
%For notational simplicity, let $U_i = (1 + X_i - \theta) / 2$, so that the $U_i$ are uniform on $[0, 1]$. First, note that we always have $\theta_n\init \ge \theta - \frac{2 n \epsilon}{m}$, because no observations will be below
% the appropriate threshold. Let $j\opt$ be the \newtext{largest} index $j$
% for which $\theta \newtext{\ge} t_{j\opt}$, and consider
%the index sets $I_{j\opt}$, $I_{j\opt + 1}$, and so on.
%The event $\theta\init_n \ge t_{j\opt} + \frac{2 n \epsilon}{m}$ may
%occur only if for each of the $n/m$ observations in the set
%$I_{j\opt + 1}$, we have $U_i \ge \frac{n \epsilon}{m}$. Thus,
%\begin{align*}
%  P_\theta\left(\theta_n\init \ge \theta + \frac{2 n \epsilon}{m}
%  \right) \le
%  P_\theta\left(\theta_n\init \ge t_{j\opt} + \frac{2 n \epsilon}{m}
%  \right)
%  \le \left(1 - \frac{n \epsilon}{m}\right)^\frac{n}{m}
%  \le \exp\left(-\frac{n^2\epsilon}{m^2}\right).
%\end{align*}
Therefore, setting the number of bins $m = \sqrt{n}$ and
the resolution $\epsilon = 2 \log n / n$, 
\begin{align}
  \label{eqn:quality-of-initial-estimate}
  \sup_{|\theta| \le \log n} P_\theta\left(|\theta_n\init - \theta| \ge
  \frac{8 \log n}{\sqrt{n}}\right) \le \frac{1}{n^4}.
\end{align}
}
%$P_\theta(\theta_n\init \ge t_{j\opt} + 4 \log n / \sqrt{n})
%\le n^{-2}$. Thus we have
%\begin{align}
%  \label{eqn:quality-of-initial-estimate}
%  \sup_{|\theta| \le \log n} P_\theta\left(|\theta_n\init - \theta| \ge
%  \frac{8 \log n}{\sqrt{n}}\right) \le \frac{1}{n^2}.
%\end{align}

The second stage estimator follows roughly the same strategy, except that
the resolution of the bins is tighter. In particular, let us assume that
$|\theta_n\init - \theta| \le \frac{8 \log n}{\sqrt{n}}$, which happens
eventually by inequality~\eqref{eqn:quality-of-initial-estimate}.  (We will
assume this tacitly for the remainder of the argument.)  Consider the
interval $\Theta_n \defeq \theta_n\init + [-\frac{16 \log n}{\sqrt{n}},
  \frac{16 \log n}{\sqrt{n}}]$ centered at $\theta_n\init$; we know that the
interval includes $[\theta - \frac{8 \log n}{\sqrt{n}}, \theta + \frac{8
    \log n}{\sqrt{n}}]$. Without loss of generality we assume $\theta_n\init
= 0$.  Following precisely the same discretization strategy as that for
$\theta_n\init$, we divide $\Theta_n$ into $m$ equal intervals, with
thresholds $t_j = -\frac{16 \log n}{\sqrt{n}} + \frac{32 (j - 1) \log n}{m
  \sqrt{n}}$; let $\epsilon_n = \frac{32 \log n}{m \sqrt{n}}$ be the width of
these intervals.  Then following exactly the same reasoning as above, we
assign indices $I_j = \{\frac{n(j - 1)}{m} + 1, \ldots, \frac{n j}{m}\}$ and
for $i \in I_j$, set $B_i = 1$ if $X_i \newtext{- 1} \ge t_j$. We define $\theta_n$ to be
the \newtext{minimal} threshold $t_j$ for which $B_i = 0$ for all observations $X_i
\in I_j$. Then following precisely the reasoning above, we have (on the
event that $|\theta_n\init - \theta| \le \frac{8 \log n}{\sqrt{n}}$)
\begin{equation*}
  P_\theta(|\theta_n - \theta|
  \ge 2 \epsilon_n)
  \le (1 - \epsilon_n)^\frac{n}{m}
  \le \exp\left(-\frac{n \epsilon_n}{m}\right)
  = \exp\left(-\frac{32 \sqrt{n} \log n}{m^2}\right).
\end{equation*}
Set $m = 4 n^{1/4}$ to obtain the claimed
result~\eqref{eqn:uniforms-are-easy}.

>>>>

\newtext{Let us now consider the probability that $\theta_n\init$ is substantially wrong. Set $\theta_M \equiv \max_{i \in I_j^*} X_i - 1$. Note that we always have $\theta\init_n \ge \theta_M$ because no observations will be above
$t_j^*+1$, and that $\theta\init_n \le \theta_M + 2n\epsilon/m$. 
 In addition, 
\begin{align*}
P_\theta\left( \left| \theta_M - \theta \right| \ge \frac{2n\epsilon}{m} \right) = \left(1 - \frac{2 n \epsilon}{m} \right)^{n/m}.
\end{align*}
Putting it all together using the triangle inequality, we have
\begin{align*}
  P_\theta\left( \left| \theta\init_n - \theta \right| \ge \frac{4n\epsilon}{m} \right) \le \left(1 - \frac{2 n \epsilon}{m} \right)^{n/m} \le e^{-2 \frac{n^2}{m^2} \epsilon}
  .
\end{align*}
%For notational simplicity, let $U_i = (1 + X_i - \theta) / 2$, so that the $U_i$ are uniform on $[0, 1]$. First, note that we always have $\theta_n\init \ge \theta - \frac{2 n \epsilon}{m}$, because no observations will be below
% the appropriate threshold. Let $j\opt$ be the \newtext{largest} index $j$
% for which $\theta \newtext{\ge} t_{j\opt}$, and consider
%the index sets $I_{j\opt}$, $I_{j\opt + 1}$, and so on.
%The event $\theta\init_n \ge t_{j\opt} + \frac{2 n \epsilon}{m}$ may
%occur only if for each of the $n/m$ observations in the set
%$I_{j\opt + 1}$, we have $U_i \ge \frac{n \epsilon}{m}$. Thus,
%\begin{align*}
%  P_\theta\left(\theta_n\init \ge \theta + \frac{2 n \epsilon}{m}
%  \right) \le
%  P_\theta\left(\theta_n\init \ge t_{j\opt} + \frac{2 n \epsilon}{m}
%  \right)
%  \le \left(1 - \frac{n \epsilon}{m}\right)^\frac{n}{m}
%  \le \exp\left(-\frac{n^2\epsilon}{m^2}\right).
%\end{align*}
Therefore, setting the number of bins $m = \sqrt{n}$ and
the resolution $\epsilon = \log n / n$, 
\begin{align}
  \label{eqn:quality-of-initial-estimate}
  \sup_{|\theta| \le \log n} P_\theta\left(|\theta_n\init - \theta| \ge
  \frac{4 \log n}{\sqrt{n}}\right) \le \frac{1}{n^2}.
\end{align}
}
%$P_\theta(\theta_n\init \ge t_{j\opt} + 4 \log n / \sqrt{n})
%\le n^{-2}$. Thus we have
%\begin{align}
%  \label{eqn:quality-of-initial-estimate}
%  \sup_{|\theta| \le \log n} P_\theta\left(|\theta_n\init - \theta| \ge
%  \frac{8 \log n}{\sqrt{n}}\right) \le \frac{1}{n^2}.
%\end{align}

The second stage estimator follows roughly the same strategy, except that
the resolution of the bins is tighter. In particular, let us assume that
$|\theta_n\init - \theta| \le \frac{8 \log n}{\sqrt{n}}$, which happens
eventually by inequality~\eqref{eqn:quality-of-initial-estimate}.  (We will
assume this tacitly for the remainder of the argument.)  Consider the
interval $\Theta_n \defeq \theta_n\init + [-\frac{16 \log n}{\sqrt{n}},
  \frac{16 \log n}{\sqrt{n}}]$ centered at $\theta_n\init$; we know that the
interval includes $[\theta - \frac{8 \log n}{\sqrt{n}}, \theta + \frac{8
    \log n}{\sqrt{n}}]$. Without loss of generality we assume $\theta_n\init
= 0$.  Following precisely the same discretization strategy as that for
$\theta_n\init$, we divide $\Theta_n$ into $m$ equal intervals, with
thresholds $t_j = -\frac{16 \log n}{\sqrt{n}} + \frac{32 (j - 1) \log n}{m
  \sqrt{n}}$; let $\epsilon_n = \frac{32 \log n}{m \sqrt{n}}$ be the width of
these intervals.  Then following exactly the same reasoning as above, we
assign indices $I_j = \{\frac{n(j - 1)}{m} + 1, \ldots, \frac{n j}{m}\}$ and
for $i \in I_j$, set $B_i = 1$ if $X_i \newtext{- 1} \ge t_j$. We define $\theta_n$ to be
the \newtext{minimal} threshold $t_j$ for which $B_i = 0$ for all observations $X_i
\in I_j$. Then following precisely the reasoning above, we have (on the
event that $|\theta_n\init - \theta| \le \frac{8 \log n}{\sqrt{n}}$)
\begin{align*}
  & P_\theta(|\theta_n - \theta|
  \ge 2 \epsilon_n)
  \le (1 - \epsilon_n)^\frac{n}{m}
  \le \exp\left(-\frac{n \epsilon_n}{m}\right) \\
  & \quad = \exp\left(-\frac{32 \sqrt{n} \log n}{m^2}\right).
\end{align*}
Set $m = 4 n^{1/4}$ to obtain the claimed
result~\eqref{eqn:uniforms-are-easy}.
